---
title: "derive_segmentation_probabilites"
author: "Sophie Su"
date: "`r Sys.Date()`"
output: html_document
---

## Document Description
This document outlines the script used to derive segmentation probabilities based on button presses from two separate studies. To calculate the segmentation probabilities, we applied Gaussian smoothing with a bandwidth of 2 seconds. Time points were selected to align with the sampling rate of the hand locations.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(magrittr)
library(lme4)
opts_chunk$set(message=FALSE, warning=FALSE)
library(stats)
#library(jtools)
library(broom)
library(arm)
library(grDevices)
library(nlme)
movie_information=data.frame(Movie=c("1.2.3_C1_trim",
                                     "2.4.1_C1_trim",
                                     "3.1.3_C1_trim",
                                     "6.3.9_C1_trim"),
                            duration=c(586.103000,
                                       645.796000,
                                       585.936000,
                                       679.429000))
#set the directory to where this file existed 
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
```

## segmentation probability 
```{r}
data_file_path <- getwd()
setwd(data_file_path)
seg_combined<-read.csv("../data/seg_combined.csv")
frame_rate=59.9 
#sampled_frame=c()
#sampled_timepoint=sampled_frame/frame_rate
seg_prob=data.frame()
hand_smoothed_grid_list_60=c("../data/movie_123_smoothed_60_by_grids.csv","../data/movie_241_smoothed_60_by_grids.csv","../data/movie_313_smoothed_60_by_grids.csv","../data/movie_639_smoothed_60_by_grids.csv") 
for(m in unique(seg_combined$Movie)){
  index=substr(gsub("\\.", "", m),1,3) 
  hand_data=read.csv(hand_smoothed_grid_list_60[grep(index,hand_smoothed_grid_list_60)])
  frame=unique(hand_data$frame)
  sampled_time=frame/frame_rate
  for (c in unique(seg_combined$Condition)){
    foo<-density(seg_combined$Sec[seg_combined$Condition==c&seg_combined$Movie==m],
                 n=round(movie_information$duration[movie_information$Movie==m]),
                 from=0, to=round(movie_information$duration[movie_information$Movie==m]),bw=2) 
    Time<-sampled_time
    sampled_density <- approx(foo$x,foo$y, xout = Time)
    scale<-length(seg_combined$Sec[seg_combined$Condition==c&seg_combined$Movie==m])/length(unique(seg_combined$SubjNum[seg_combined$Condition==c&seg_combined$Movie==m]))
    temp<-data.frame(Time=Time,Segmentation_Prob=scale*sampled_density$y)
    temp$Movie=m
    temp$Condition=c
    seg_prob=rbind(seg_prob,temp)
  }
}
```



## visualization for publication
```{r}
seg_prob %>%
    mutate(Movie = case_when(
        Movie == "1.2.3_C1_trim" ~ "Exercise",
        Movie == "2.4.1_C1_trim" ~ "Grooming",
        Movie == "3.1.3_C1_trim" ~ "Breakfast",
        Movie == "6.3.9_C1_trim" ~ "Cleaning",
        TRUE ~ Movie  # Keep other values unchanged
    )) %>%
    ggplot(aes(x = Time, y = Segmentation_Prob, color = Condition)) +
    facet_wrap(~ Movie, ncol = 2) +
    geom_line() +
    labs(x = "Time (s)", y = "Segmentation Probability") +
    scale_color_manual(values = c("coarse" = "red", "fine" = "blue")) +
    theme_classic() +  # Use a minimal theme
    annotate("text", x = 460, y = 0.16, label = "coarse", color = "red", hjust = 1, vjust = 1, size = 5) +
    annotate("text", x = 460, y = 0.2, label = "fine", color = "blue", hjust = 1, vjust = 1, size = 5)+theme(legend.position = "none")+theme(text = element_text(size = 20),
        plot.background = element_rect(color = "white"),
        aspect.ratio = 1)
##ggsave("Seg_Prob_Movie.tiff", units="in", width=7, height=8, dpi=300)
```

### Supplemental Material: Compare two distinct populations
```{r}
# Initialize an empty data frame to store the results
seg_prob <- data.frame()

# Loop over each unique movie
for(m in unique(seg_combined$Movie)){
  # Clean the movie names to create indices
  index = substr(gsub("\\.", "", m), 1, 3)
  
  # Read hand data based on the index
  hand_data = read.csv(hand_smoothed_grid_list_60[grep(index, hand_smoothed_grid_list_60)])
  frame = unique(hand_data$frame)
  sampled_time = frame / frame_rate
  
  # Loop over each unique dataset
  for (d in unique(seg_combined$dataset)) {
    # Loop over each unique condition within the dataset
    for (c in unique(seg_combined$Condition[seg_combined$dataset == d])) {
      # Calculate the density for the current condition, movie, and dataset
      foo <- density(seg_combined$Sec[seg_combined$Condition == c & seg_combined$Movie == m & seg_combined$dataset == d],
                     n = round(movie_information$duration[movie_information$Movie == m]),
                     from = 0, to = round(movie_information$duration[movie_information$Movie == m]), bw = 2)
      Time <- sampled_time
      sampled_density <- approx(foo$x, foo$y, xout = Time)
      
      # Calculate scale factor
      scale <- length(seg_combined$Sec[seg_combined$Condition == c & seg_combined$Movie == m & seg_combined$dataset == d]) /
               length(unique(seg_combined$SubjNum[seg_combined$Condition == c & seg_combined$Movie == m & seg_combined$dataset == d]))
      
      # Create a temporary data frame to store results
      temp <- data.frame(Time = Time, Segmentation_Prob = scale * sampled_density$y, Movie = m, Condition = c, Dataset = d)
      
      # Bind the results to the main results data frame
      seg_prob <- rbind(seg_prob, temp)
    }
  }
}

seg_prob %>%
    mutate(Movie = case_when(
        Movie == "1.2.3_C1_trim" ~ "Exercise",
        Movie == "2.4.1_C1_trim" ~ "Grooming",
        Movie == "3.1.3_C1_trim" ~ "Breakfast",
        Movie == "6.3.9_C1_trim" ~ "Cleaning",
        TRUE ~ Movie  # Keep other values unchanged
    )) %>%
  mutate(Combined = paste(Movie, Condition, sep = "-"))%>%
    ggplot(aes(x = Time, y = Segmentation_Prob, color = Dataset)) +
    facet_wrap(~ Combined, ncol = 4) +
    geom_line() +
    labs(x = "Time (s)", y = "Segmentation Probability") +
    scale_color_manual(values = c("first_dataset" = "lightblue", "second_dataset" = "darkblue")) +
    theme_classic() +  # Use a minimal theme
    annotate("text", x = 460, y = 0.4, label = "Online Participants", color = "lightblue", hjust = 1, vjust = 1, size = 2) +
    annotate("text", x = 460, y = 0.5, label = "fMRI Participants", color = "darkblue", hjust = 1, vjust = 1, size = 2)+theme(legend.position = "none")+theme(text = element_text(size = 11),
        plot.background = element_rect(color = "white"),
        aspect.ratio = 1)
##ggsave("Seg_Prob_Movie_twodataset.tiff", units="in", width=14, height=8, dpi=300)

```
