---
title: "Predictive_Looking_Analysis"
author: "Sophie Su"
date: "`r Sys.Date()`"
output: html_document
---

# Predictive Modeling Errors

This R Markdown file contains the code to analyze predictive modeling errors and reproduce the figures in the manuscript.

### Read in the data and relevant pacakges 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(tidyverse)
library(magrittr)
library(lme4)
opts_chunk$set(message=FALSE, warning=FALSE)
library(stats)
#library(jtools)
library(broom)
library(arm)
library(grDevices)
library(nlme)
data_file_path <- getwd()
setwd(data_file_path)
cl1 <- glmerControl(optimizer = "bobyqa", calc.derivs = FALSE,
    optCtrl = list(maxfun = 1e+09), check.conv.grad = .makeCC("warning",
        tol = 0.002, relTol = NULL), check.conv.singular = .makeCC(action = "message",
        tol = 1e-09), check.conv.hess = .makeCC(action = "warning",
        tol = 0.000001))
c1=lmeControl(maxIter=100000, msMaxIter=100000, niterEM=300, msMaxEval=100000, tolerance=.000001,msTol=.0000001,returnObject=TRUE,minAbsParApVar=.05, opt = c("nlminb"), optimMethod="BFGS")
#read into the dataframe  
all_movie<-read.csv("../data/all_movie.csv", header = TRUE, sep = ",")

```

## Stepwise model comparsion approach to examine predictive looking.
```{r}
m15=glmer(hand_bino ~gaze_value + gaze_past1+gaze_past2+gaze_past3+gaze_past4+gaze_past5+gaze_past6+gaze_past7+gaze_past8+gaze_past9+gaze_past10+gaze_past11+gaze_past12+gaze_past13+gaze_past14+(1|grid)+(1|movie),data=all_movie,family = "binomial"("logit"),control=cl1)
model_data=na.omit(all_movie[ , c(all.vars(formula(m15)),"frame")])
m14=glmer(hand_bino ~gaze_value + gaze_past1+gaze_past2+gaze_past3+gaze_past4+gaze_past5+gaze_past6+gaze_past7+gaze_past8+gaze_past9+gaze_past10+gaze_past11+gaze_past12+gaze_past13+(1|grid)+(1|movie),data=model_data,family = "binomial"("logit"),control=cl1)
m13=glmer(hand_bino ~gaze_value + gaze_past1+gaze_past2+gaze_past3+gaze_past4+gaze_past5+gaze_past6+gaze_past7+gaze_past8+gaze_past9+gaze_past10+gaze_past11+gaze_past12+(1|grid)+(1|movie),data=model_data,family = "binomial"("logit"),control=cl1)
m12=glmer(hand_bino ~gaze_value + gaze_past1+gaze_past2+gaze_past3+gaze_past4+gaze_past5+gaze_past6+gaze_past7+gaze_past8+gaze_past9+gaze_past10+gaze_past11+(1|grid)+(1|movie),data=model_data,family = "binomial"("logit"),control=cl1)
m11=glmer(hand_bino ~gaze_value + gaze_past1+gaze_past2+gaze_past3+gaze_past4+gaze_past5+gaze_past6+gaze_past7+gaze_past8+gaze_past9+gaze_past10+(1|grid)+(1|movie),data=model_data,family = "binomial"("logit"),control=cl1)
m10=glmer(hand_bino ~gaze_value + gaze_past1+gaze_past2+gaze_past3+gaze_past4+gaze_past5+gaze_past6+gaze_past7+gaze_past8+gaze_past9+(1|grid)+(1|movie),data=model_data,family = "binomial"("logit"),control=cl1)
m9=glmer(hand_bino ~gaze_value + gaze_past1+gaze_past2+gaze_past3+gaze_past4+gaze_past5+gaze_past6+gaze_past7+gaze_past8+(1|grid)+(1|movie),data=model_data,family = "binomial"("logit"),control=cl1)
m8=glmer(hand_bino ~gaze_value + gaze_past1+gaze_past2+gaze_past3+gaze_past4+gaze_past5+gaze_past6+gaze_past7+(1|grid)+(1|movie),data=model_data,,family = "binomial"("logit"),control=cl1)
m7=glmer(hand_bino ~gaze_value + gaze_past1+gaze_past2+gaze_past3+gaze_past4+gaze_past5+gaze_past6+(1|grid)+(1|movie),data=model_data,,family = "binomial"("logit"),control=cl1)
#model_data=na.omit(movie1.2.3[ , c(all.vars(formula(m7)),"frame")])
m1=glmer(hand_bino ~gaze_value+(1|grid)+(1|movie),data=model_data,family = "binomial"("logit"),control=cl1)
m2=glmer(hand_bino ~gaze_value + gaze_past1+(1|grid)+(1|movie),data=model_data,family = "binomial"("logit"),control=cl1)
m3=glmer(hand_bino ~gaze_value + gaze_past1+gaze_past2+(1|grid)+(1|movie),data=model_data,family = "binomial"("logit"),control=cl1)
m4=glmer(hand_bino ~gaze_value + gaze_past1+gaze_past2+gaze_past3+(1|grid)+(1|movie),data=model_data,family = "binomial"("logit"),control=cl1)
m5=glmer(hand_bino ~gaze_value + gaze_past1+gaze_past2+gaze_past3+gaze_past4+(1|grid)+(1|movie),data=model_data,family = "binomial"("logit"),control=cl1)
m6=glmer(hand_bino ~gaze_value + gaze_past1+gaze_past2+gaze_past3+gaze_past4+gaze_past5+(1|grid)+(1|movie),data=model_data,family = "binomial"("logit"),control=cl1)
m15=glmer(hand_bino ~gaze_value + gaze_past1+gaze_past2+gaze_past3+gaze_past4+gaze_past5+gaze_past6+gaze_past7+gaze_past8+gaze_past9+gaze_past10+gaze_past11+gaze_past12+gaze_past13+gaze_past14+(1|grid)+(1|movie),data=model_data,family = "binomial"("logit"),control=cl1)
AIC=anova(m1,m2,m3,m4,m5,m6,m7,m8,m9,m10,m11,m12,m13,m14,m15)$AIC 
BIC=anova(m1,m2,m3,m4,m5,m6,m7,m8,m9,m10,m11,m12,m13,m14,m15)$BIC
graph_data=data.frame(model=c("0s","1s","2s","3s","4s","5s","6s","7s","8s","9s","10s","11s","12s","13s","14s"),AIC=AIC, BIC=BIC)
visual_data<- pivot_longer(graph_data, cols = c("AIC", "BIC"), 
                        names_to = "Parameter", values_to = "Value")
visual_data$model <- factor(visual_data$model, levels = c("0s", "1s", "2s", "3s", "4s", "5s", "6s", "7s", "8s", "9s", "10s", "11s", "12s", "13s", "14s"))


ggplot(data=visual_data, aes(x=model, y=Value,color=Parameter,group=Parameter))+geom_point()+geom_line(size=2)+scale_color_manual(values = c("BIC" = "green", "AIC" = "black"),guide = "none") +
labs(x = "Number of previous seconds of eye data in the predictive model")+annotate("text", x = 5, y = 17775, label = "BIC", color = "green")+annotate("text", x = 5, y = 17700, label = "AIC", color = "black")+theme(text = element_text(size = 15),plot.background = element_rect(color = "white"),aspect.ratio = 1)+theme(legend.position = "none")+theme_classic()
##ggsave("model_fit.tiff", units="in", width=7, height=8, dpi=300)
```

#Predicitve Looking Error and Segmentation Probabilities 
```{r}
modeling_10=glmer(hand_bino ~gaze_value + gaze_past1+gaze_past2+gaze_past3+gaze_past4+gaze_past5+gaze_past6+gaze_past7+gaze_past8+gaze_past9+(1|grid)+(1|movie),data=all_movie,family = "binomial"("logit"),control=cl1)
modeling_10_data=na.omit(all_movie[ , c(all.vars(formula(modeling_10)),"frame")])
dev_m10<-abs(resid(modeling_10, type="deviance"))
pear_m10<-abs(resid(modeling_10, type="pearson"))
work_m10<-abs(resid(modeling_10, type="response"))

PE_data_m10<-data.frame(frame=modeling_10_data$frame,grid=modeling_10_data$grid,eid=paste(modeling_10_data$frame,modeling_10_data$movie),
                    dev=dev_m10,
                    pear=pear_m10,
                    work=work_m10)
PE_data_m10<-PE_data_m10%>%
  mutate(Movie=str_sub(eid,-5,-1))%>%
 mutate(frame=as.numeric(str_sub(eid, start=1,end=-7)))
```

## Segmentation Probabilities
```{r}
frame_rate=59.9
seg_des<-read.csv("../data/segmentation_probability_sampled.csv")
seg_des %<>% 
  #dplyr::filter(Movie=="1.2.3_C1_trim")%>%
  mutate(frame=round(Time*frame_rate))%>%
  mutate(Movie=gsub("_C1_trim", "", Movie))%>%
  mutate(eid=paste(frame, Movie))%>%
  rename(probability=Segmentation_Prob)%>%
  dplyr::select("frame", "probability","Condition","eid","Movie")
```

# plot the relationship between predicitve looking error and segmentation probabilites
```{r}
combined_model_10<-left_join(PE_data_m10, seg_des, by="eid")
combined_model_10 <- na.omit(combined_model_10)
model1<-lmer(probability ~dev*Condition+(1|grid)+(1|Movie.x)+(1|frame.x),data=combined_model_10,control = lmerControl(optimizer = "bobyqa"))

summary(model1)
combined_model_10$estimated_prob=predict(model1)

combined_model_10_scaled<-combined_model_10%>%
  group_by(Movie.x)%>%
  mutate(scaled_dev=scale(dev))%>%
  ungroup()%>%
  group_by(Movie.x, Condition)%>%
  mutate(scaled_est_prob=scale(estimated_prob))
combined_model_10_scaled%>%
  group_by(Movie.x)%>%
  summarise(mean_dev=mean(scaled_est_prob, na.rm=TRUE))
combined_model_10_scaled <- combined_model_10_scaled %>%
  mutate(Movie = dplyr::recode(Movie.x,
                          "1.2.3" = "Exercise",
                          "2.4.1" = "Grooming",
                          "3.1.3" = "Breakfast",
                          "6.3.9" = "Cleaning"))

combined_model_10_scaled%>%
ggplot(aes(x=scaled_dev, y = scaled_est_prob,color=Condition))+geom_smooth(method="lm",se=TRUE)+facet_grid( ~ Movie)+labs(x="Scaled Predictive Looking Error", y="Estimated Segmentation Probability")+
  scale_y_continuous(name="Estimated Segmentation Probability",breaks=seq(-1,1,1)) +  # Use a minimal theme
    annotate("text", x = 3, y = 0.018, label = "coarse", color = "red", hjust = 1, vjust = 1, size = 3) +
    annotate("text", x = 3, y = 0.040, label = "fine", color = "blue", hjust = 1, vjust = 1, size = 3)+theme_classic()+theme(legend.position = "none")
  
  
combined_model_10_scaled %>%
    group_by(Condition, Movie.x) %>%
    mutate(scaled_est_prob_new = scale(estimated_prob)) %>%
    ggplot(aes(x = scaled_dev, y = scaled_est_prob_new, color = Condition)) +
    geom_smooth(method = "lm", se = TRUE) +
    facet_wrap(~ Movie, nrow = 2, scales = "fixed") +  # Use facet_wrap with nrow set to 2
    theme(aspect.ratio = 1)+labs(x="Scaled Predictive Looking Error", y="Scaled Estimated Segmentation Probability")+
  scale_color_manual(values = c("coarse" = "red", "fine" = "blue"))+
  theme_classic() +  # Use a minimal theme
    annotate("text", x = 4.5, y = 0.3, label = "coarse", color = "red", hjust = 1, vjust = 1, size=5) +
    annotate("text", x = 4.5, y = 0.4, label = "fine", color = "blue", hjust = 1, vjust = 1, size=5)+theme(legend.position = "none")+
  theme(text = element_text(size = 20),
        plot.background = element_rect(color = "white"),
        aspect.ratio = 1)

##ggsave("PE_Seg_Prob.tiff", units="in", width=7, height=8, dpi=300)
```

### Cross correlation between Predictive Looking Error and Segmentation Probabilities
```{r} 
cross_correlation_results<-data.frame()
combined_model_10_cross_compare<-combined_model_10%>%
  mutate(Movie.x = case_when(
    Movie.x == "1.2.3" ~ "Exercise",
    Movie.x == "2.4.1" ~ "Grooming",
    Movie.x == "3.1.3" ~ "Breakfast",
    Movie.x == "6.3.9"~ "Cleaning",
    TRUE ~ "Other Movie" # TRUE here means for all other cases
  ))
for(m in unique(combined_model_10_cross_compare$Movie.x)){
  for(i in c("coarse","fine")){
    d=str_to_title(i)
    dev=na.omit(combined_model_10_cross_compare$dev[combined_model_10_cross_compare$Movie.x==m&combined_model_10_cross_compare$Condition==i])
    probability=na.omit(combined_model_10_cross_compare$probability[combined_model_10_cross_compare$Movie.x==m&combined_model_10_cross_compare$Condition==i])
    png(paste0(data_file_path,"cross_cor_movie",m,"condition", i, ".png"),width = 800, height = 600)
    ccf_result <- ccf(dev, probability, lag.max = NULL, plot =  TRUE, main = paste0(m,':',d," Segmentation"))
    #plot()
    dev.off()
    temp=data.frame(movie=m, condition=i, max_corr_lag =ccf_result$lag[which.max(ccf_result$acf)])
    cross_correlation_results<-rbind(cross_correlation_results, temp)
  }
}
#print(cross_correlation_results)
##write.csv(cross_correlation_results, "cross_correlation_results.csv")
```




## Updating Signals VS predictive looking errors 
```{r}
PE_eye_uncertainty=read.csv("../data/PE_eye_uncertainty_cleaned.csv")%>%
  dplyr::select("eid","dev", "Movie.x", "frame.x", "sem_uncertainty", "grid")%>%
  rename(Movie=Movie.x,
         frame=frame.x, 
         eye_PE=dev
         )%>%
  mutate(eid=paste(eid, grid))
# dataset with SEM's prediction error and predictive looking error
PE_eye_SEM=read.csv("../data/PE_eye_SEM_cleaned.csv")%>%
  dplyr::select("eid","dev", "Movie.x", "frame.x", "sem_pe", "grid")%>%
  rename(Movie=Movie.x,
         frame=frame.x, 
         eye_PE=dev)%>%
  mutate(eid=paste(eid,grid ))

eye_SEM<-left_join(PE_eye_SEM, PE_eye_uncertainty, by="eid")%>%
  dplyr::select("eid","eye_PE.x","sem_uncertainty", "Movie.x", "frame.x", "sem_pe", "grid.x")%>%
  rename(Movie=Movie.x,
         frame=frame.x, 
         eye_PE=eye_PE.x,
         grid=grid.x)
names(eye_SEM)
eye_SEM<-eye_SEM%>%
  group_by(Movie)%>%
  mutate(sem_pe_scaled=scale(sem_pe),
         eye_PE_scaled=scale(eye_PE), 
         sem_uncertainty_scaled=scale(sem_uncertainty))
eye_SEM_visual<-eye_SEM%>%
  dplyr::select(sem_pe_scaled,eye_PE_scaled, sem_uncertainty_scaled, Movie )
eye_SEM_visual %>%
  mutate(Movie = case_when(
   Movie == "1.2.3" ~ "Exercise",
    Movie == "2.4.1" ~ "Grooming",
    Movie == "3.1.3" ~ "Breakfast",
   Movie=="6.3.9" ~ "Cleaning",
    TRUE ~ Movie  # Keep other values unchanged
  ))%>%
  pivot_longer(cols = c(sem_pe_scaled, sem_uncertainty_scaled),
               names_to = "signals",
               values_to = "value")%>%
  ggplot(aes(x=eye_PE_scaled, y=value, color = signals))+
  geom_smooth(method="lm")+facet_wrap(~ Movie, nrow = 2, scales = "fixed")+theme(aspect.ratio = 1)+labs(x="Scaled Predictive Looking Error", y="Scaled Model Updating Signals")+
  scale_color_manual(values = c("sem_pe_scaled" = "darkred", "sem_uncertainty_scaled" = "purple")) +
  theme_classic() +  # Use a minimal theme
   annotate("text", x = 9, y = 0.8, label = "Prediction Error", color = "darkred", hjust = 1, vjust = 1, size = 5) +
    annotate("text", x = 9, y = 0.7, label = "Prediction Uncertainty", color = "purple", hjust = 1, vjust = 1, size = 5)+theme(legend.position = "none")+
  theme(text = element_text(size = 20),
        plot.background = element_rect(color = "white"),
        aspect.ratio = 1)

##ggsave("Signals_Seg_Prob.tiff", units="in", width=7, height=8, dpi=300)
```

## Statistical Comparison between Predictive Looking Error and Other SEMs' Updating Mechanisms.

```{r}
eye_SEM<-eye_SEM%>%
  group_by(Movie)%>%
  mutate(eye_PE_sd=scale(eye_PE), 
         sem_pe_sd=scale(sem_pe), 
         sem_uncertainty_sd=scale(sem_uncertainty))
model1<-lme(eye_PE_sd ~ sem_pe_sd, random=~ 1|Movie/grid, data=eye_SEM)
summary(model1)
model2<-lme(eye_PE_sd ~ sem_uncertainty_sd,  random=~ 1|Movie/grid, data=eye_SEM)
summary(model2)
anova(model1, model2)
```